### 2025-01-28 ESC195 Lecture Notes/main.tex
jan 28

$$
\int_{0}^{2} \sqrt{x} \, dx = \lim_{n \to \infty} \sum_{i=1}^n \sqrt{x_i} \Delta x_i
$$

where:
\begin{itemize}
    \item $x_0 = 0, \, x_n = 2$,
    \item $x_i = i^2 \cdot \frac{2}{n^2}$ for $i = 0, 1, 2, \dots, n$,
    \item $\Delta x_i = x_i - x_{i-1} = \frac{2}{n^2} \cdot (2i - 1)$.
\end{itemize}

@@

\subsection{Integration By Parts}
Using the product rule:
$$
\frac{d}{dx} \big[ u(x) v(x) \big] = u'(x)v(x) + u(x)v'(x),
$$
integrating both sides with respect to $x$ gives:
$$
u(x)v(x) = \int u'(x) v(x) \, dx + \int u(x) v'(x) \, dx.
$$
Rearranging this:
$$
\int u'(x) v(x) \, dx = u(x)v(x) - \int u(x) v'(x) \, dx.
$$

@@

\appendix
\input{appendix.tex}
\end{document}

>>>>

$$\int_{0}^{2} \sqrt{x} \, dx = \lim_{n \to \infty} \sum_{i=1}^n \sqrt{x_i} \Delta x_i$$
, where:
\begin{itemize}
    \item $x_0 = 0, \, x_n = 2$,
    \item $x_i = i^2 \cdot \frac{2}{n^2}$ for $i = 0, 1, 2, \dots, n$,
    \item $\Delta x_i = x_i - x_{i-1} = \frac{2}{n^2} \cdot (2i - 1)$.
\end{itemize}

@@

\subsection{Integration By Parts}
Using the product rule:
$$
\frac{d}{dx} \big[ u(x) v(x) \big] = u'(x)v(x) + u(x)v'(x),
$$
integrating both sides with respect to $x$ gives:
$$
u(x)v(x) = \int u'(x) v(x) \, dx + \int u(x) v'(x) \, dx.
$$
Rearranging this:
$$\int u'(x) v(x) \, dx = u(x)v(x) - \int u(x) v'(x) \, dx.$$

@@

\section{Polar Coordinates}
\paragraph{Polar Coordinates} Given a point $P$ in the plane, we can define the polar coordinates of $P$ as $(r, \theta)$, where $r$ is the distance from the origin to $P$ and $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to $P$.
\paragraph{Transformation} We can convert between polar and Cartesian coordinates as follows:
    x = r\cos \theta, \quad y = r\sin \theta
\paragraph{Reverse Transformation} We can convert between Cartesian and polar coordinates as follows:
    r = \sqrt{x^2 + y^2}, \quad \theta = \arctan\left(\frac{y}{x}\right)
\paragraph{Note} The angle $\theta$ is not unique, as $\theta + 2\pi n$ for $n \in \mathbb{Z}$ represents the same point, beware of this when converting between polar and Cartesian coordinates.
\begin{example}[Lines]
        y = mx + b &\implies \theta = \alpha = \arctan m \\
        x = a &\implies r = a\sec \theta \\
        y = a &\implies r = a\csc \theta
\begin{example}[Cirlces]
        x^2 + y^2 = a^2 &\implies r = a \\
    Let $r = 6\sin \theta$, we have:
        r^2 &= 6r\sin \theta \\
        x^2 + y^2 &= 6y \\
        x^2 + (y-3)^2 &= 9
    We deduce that the curve is a circle with radius $3$ and center $(0,3)$.
\paragraph{Cylinrical Coordinates} Given a point $P$ in space, we can define the cylindrical coordinates of $P$ as $(r, \theta, z)$, where $r$ is the distance from the $z$-axis to $P$, $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to $P$ onto the $xy$-plane, and $z$ is the distance from the $xy$-plane to $P$.
\paragraph{Spherical Coordinates} Given a point $P$ in space, we can define the spherical coordinates of $P$ as $(\rho, \theta, \phi)$, where $\rho$ is the distance from the origin to $P$, $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to $P$ onto the $xy$-plane, and $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to $P$.
\subsection{Graphing in Polar Coordinates}
    Let $r = \frac{1}{2} + \cos \theta$, we first figure out the origin of the curve by setting $r = 0$:
        0 &= \frac{1}{2} + \cos \theta \\
        \theta &= \frac{2\pi}{3},\, \frac{4\pi}{3}
    We then deduce the maximum and minimum of the curve to the origin by setting $\frac{dr}{d\theta} = 0$:
        \frac{dr}{d\theta} &= -\sin \theta = 0 \\
        \theta &= 0,\, \pi,\, 2\pi
    We also look for symmetry in the curve by checking $r(\theta) = r(-\theta)$:
        \frac{1}{2} + \cos \theta &= \frac{1}{2} + \cos(-\theta) \\
    Thus, the curve is symmetric about the $x$-axis. We also look for symmetry about the $y$-axis by checking $r(\theta) = r(\pi + \theta)$:
        \frac{1}{2} + \cos \theta \neq \frac{1}{2} + \cos(\pi + \theta) \\
    Thus, the curve is not symmetric about the $y$-axis. We also look for intervals in which the curve is increasing or decreasing by checking $r'(\theta) > 0$ or $r'(\theta) < 0$.
\paragraph{Common Polar Curves} Some common polar curves are:
    \item \textbf{Cardioid}: $r = a(1 + \cos \theta)$ (heart-shaped)
    \item \textbf{Circle}: $r = a\cos \theta$ 
    \item \textbf{Limascons}: $r = a + b\sin \theta$ (heart with a hole)
    \item \textbf{Lemniscate}: $r^2 = a^2\sin 2\theta$ (infinity symbol, diagonal) \vspace{1em} $\quad r^2 = a^2\cos 2\theta$ (infinity symbol, horizontal)
    \item \textbf{Petal Curves}: $r = a\cos n\theta$ ($mn$-petal flower) \vspace{1em} $\quad r = a\sin n\theta$ ($mn$-petal flower); $m=2$ if $n$ is even, $m=1$ if $n$ is odd.
\subsection{The intersection of Polar Curves}
    Let $r_1 = \sin \theta$, $r_2 = -\cos \theta$, we have:
        \sin \theta &= -\cos \theta \\
        \tan \theta &= -1 \\
        \theta &= \frac{3\pi}{4},\, \frac{7\pi}{4}
    At these points, we can find the $x$ and $y$ coordinates of the intersection points. But also, we must check for the origin point $(x,y) = (0,0)$. Also $r_1 = r_2$ is not always reliable so we have to check it intuitively.
\appendix
\input{appendix.tex}
\end{document}

### 2025-01-28 MAT185 Lecture Notes/main.tex
jan 28

\subsection{Linear combinations}
\begin{definition}[Linear Combination]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
    \end{align*}
    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\begin{definition}[Linear independence]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If and only if the equation:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
    \end{align*}
    has \textbf{only} the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.

    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
\end{definition}

\begin{definition}[Span]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
    \begin{align}
        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
    \end{align}
    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\begin{definition}[Basis]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$. If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent and $\text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = V$, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ is a basis of $V$.
\end{definition}
\end{document}

>>>>

\section{Linear combinations}
\begin{definition}[Linear Combination]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
    \end{align*}
    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\subsection{Linear Independence, Span, and Basis}
\begin{definition}[Linear independence]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If and only if the equation:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
    \end{align*}
    has \textbf{only} the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.

    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
\end{definition}
\begin{definition}[Span]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
    \begin{align}
        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
    \end{align}
    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\begin{theorem}[Unique Representation]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}/0$, and that they span $V$. For any $\textbf{v}_{n+1} \in V$, if $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent, then the equation:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{v}_{n+1}  
    \end{align*} 
    is the only way to express $\textbf{v}_{n+1}$ as a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$.
\end{theorem}
We can prove by contrapositive:
\begin{proof}
    Assume that there exists another way to express $\textbf{v}_{n+1}$ as a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$, then:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \beta_1 \textbf{v}_1 + \beta_2 \textbf{v}_2 + \ldots + \beta_n \textbf{v}_n
    \end{align*}
    Rearranging the equation, we have:
    \begin{align*}
        (\alpha_1 - \beta_1) \textbf{v}_1 + (\alpha_2 - \beta_2) \textbf{v}_2 + \ldots + (\alpha_n - \beta_n) \textbf{v}_n = \textbf{0}
    \end{align*}
    Since ($\alpha_i - \beta_i$) for $i = 1, 2, \ldots, n$ are not all zero, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly dependent. Thus, the contrapositive is true.
\end{proof}
\begin{theorem}[Growing and Pruning] The following are the two theorems describe the relationship between linear independence and the span: \\ \\
    \textbf{Growing} Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ be linearly independent. If $\textbf{v}_{n+1} \in V$ and $\textbf{v}_{n+1} \notin \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n)$, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n, \textbf{v}_{n+1}$ are linearly independent. \\ \\
    \textbf{Pruning} Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$. If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly dependent, then there exists a $\textbf{v}_i$ such that $\textbf{v}_i \in \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_{i-1}, \textbf{v}_{i+1}, \ldots, \textbf{v}_n)$.
\end{theorem}
\begin{definition}[Basis]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$. If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent and $\text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = V$, then $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ is a basis of $V$.
\end{definition}
\subsection{Fundamental Theorem of Linear Algebra}
\begin{theorem}[Fundamental Theorem of Linear Algebra]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$. If $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are linearly independent and $\text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = V$, then:
    \begin{enumerate}
        \item Any basis of $V$ has the same number of elements.
        \item Any linearly independent set of $V$ has at most $n$ elements.
        \item Any spanning set of $V$ has at least $n$ elements.
    \end{enumerate}
\end{theorem}
\end{document}

### 2025-01-28 MSE160 Lecture Notes/main.tex
jan 28

@@

@@

@@

\paragraph{Simple Cubic} The simple cubic structure has the following properties:

@@

    \item \textbf{Packing Density} Rare due to low packing density (only known example: Polonium, Po)

@@

    \item \textbf{Coordination Number} Number of nearest neighbor

@@

>>>>

\subsection{Material Performance}

@@

\section{Atomic Structures}
\subsection{General Definitions}

@@

\begin{equation} \label{eq:theoretical-density}

@@

\subsection{Describing Basic Unit Cells}
\begin{definition}[Coordination Number (CN)]
    The coordination number is the number of nearest neighbors an atom has in a crystal structure.
\begin{definition}[Atomic Packing Factor (APF)]
    The atomic packing factor is the fraction of the volume of the unit cell that is occupied by hard spheres. It is given by:
        \text{APF} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell}}
\paragraph{Simple Cubic (SC)} The simple cubic structure has the following properties:

@@

    \item \textbf{Rare Packing Density}  Due to low packing density (only known example: Polonium, Po)

@@

    \item \textbf{Coordination Number} Number of nearest neighbor: 6

@@

\paragraph{Face-Centered Cubic (FCC)} The face-centered cubic structure has an atom on the centre of each face of the cube. It has the following properties:
    \item \textbf{Ductile} and \textbf{Malleable} due to the close-packed planes. Common examples include: \\ $\ch{Al}\, ,\ch{Cu}\, ,\ch{Au}\, ,\ch{Ag}$.
    \item \textbf{Coordination Number} Number of nearest neighbor: 12
    \item \textbf{Close-packed directions} As cube diagonals.
    \item \textbf{Theortical Density} The theoretical density of FCC is derived from Equation (\ref{eq:theoretical-density}):
    \begin{align}
        \rho &= \frac{nM}{V_c N_A} \nonumber \\
        \intertext{With $n = 4$ (8 corners $\times \frac{1}{8}$ atom each + 6 faces $\times \frac{1}{2}$ atom each) and $V_c = a^3$ (where $a$ is the \textbf{lattice parameter}):}
        \rho_\text{FCC} &= \frac{4M}{a^3 N_A} \nonumber \\
        \intertext{We can then obtain $a$ via the relationship between the atomic radius, $r$, and the lattice parameter via the geometry on the close-packed direction (face diagonal):}
        a &= 2\sqrt{2}r \nonumber \\
        \intertext{Substitute $a$ back into the equation:}
        \rho_\text{FCC} &= \frac{4M}{(2\sqrt{2}r)^3 N_A} = \frac{4M}{16\sqrt{2}r^3 N_A} \nonumber \\
        &= \frac{M}{4\sqrt{2}r^3 N_A} 
    \end{align}
    \item \textbf{Atomic Packing Factor} The atomic packing factor of FCC is given by:
        \text{APF}_\text{FCC} = \frac{\text{Volume of atoms in unit cell}}{\text{Volume of unit cell} } = \frac{4 \times \frac{4}{3}\pi r^3}{a^3} = \frac{4 \times \frac{4}{3}\pi r^3}{(2\sqrt{2}r)^3} = \frac{\pi}{6\sqrt{2}} = \textbf{0.74}
    This is the highest possible packing factor for spheres. Thus, no other structure can have a higher packing factor.
\begin{example}[Density of Aluminum]
    Given that the atomic radius of aluminum is $r = 143$ pm and the atomic mass is $M = 26.98$ g/mol, we can calculate the density of aluminum with FCC structure. We have:
        \rho_\text{Al} &= \frac{M}{4\sqrt{2}r^3 N_A} = \frac{26.98}{4\sqrt{2} \times (143 \times 10^{-12})^3 \times 6.022 \times 10^{23}} \\
        &= \frac{26.98}{4\sqrt{2} \times (143 \times 10^{-12})^3 \times 6.022 \times 10^{23}} = 2.7 \text{ g/cm}^3
\section{Geometric Descriptions of Crystals}
\subsection{Crystallographic Directions and Planes}
\paragraph{Point Coordinates} A lattice positon in a unit cell is determined as fractional multiples of the unit cell edge lengths $a$, $b$, and $c$. (Defining the origin at one corner of the unit cell and $a=b=c=\textbf{1}$)
\paragraph{Crystallographic Directions} A crystallographic direction is a line between two points in a crystal. It is denoted by square brackets, $[uvw]$. The direction is determined from the ``tail'' to the ``head'' of the vector. Also:
    \item It \textbf{shall not} include commas; and
    \item It \textbf{shall be reduced}to the smallest integers; and
    \item Negative signs are represented by a bar over the number.
\begin{example}
    The crystallographic direction from $(0,0,0)$ to $(-1,0,1/2)$ is $[\bar{2}01]$.
\paragraph{Family of Directions} A family of directions is a set of directions that are parallel to each other. It is denoted by $<uvw>$.
\begin{example}
    The family of directions $<100>$ is the set of directions $[100]$, $[010]$, $[001]$, $[\bar{1}00]$, $[0\bar{1}0]$, and $[00\bar{1}]$.
\paragraph{Angle between Directions} The angle between two directions in a \textbf{cubic} crystal is given by:
    \cos \theta = \frac{h_1h_2 + k_1k_2 + l_1l_2}{\sqrt{h_1^2 + k_1^2 + l_1^2}\sqrt{h_2^2 + k_2^2 + l_2^2}}
, which is simply the normalized dot product of the two vectors.
\paragraph{Crystallographic Planes} A crystallographic plane is a set of parallel planes in a crystal. It is denoted by $(hkl)$. The plane is determined by the intercepts on the $x$, $y$, and $z$ axes. 
\paragraph{} Generally, we find the intercepts by finding the points where the plane intersects the axes. We then take the reciprocals of the intercepts and multiply by the smallest integer to get the Miller Indices.
\paragraph{Family of Planes} A family of planes is a set of planes that are parallel to each other. It is denoted by $\{hkl\}$.
\paragraph{Distance between Planes} The distance between two planes with miller indices $(h_1k_1l_1)$ and $(h_2k_2l_2)$ is given by:
    d = \frac{a}{\sqrt{h^2 + k^2 + l^2}}
, where $a$ is the lattice parameter.
